---
title: "Untitled"
output: html_document
date: "2023-02-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in the data set and remove all of the variables with zero variance (have only a single value).  Also drop all of the variables that will be used as responses in modeling.

```{r load-libs-and-data, warning = FALSE}
library(tidyverse)

anes <- read_csv("data/raw/anes_timeseries_2020_csv_20220210.csv",
                 show_col_types = FALSE)

# Keep only pre- and post-election vars and pre- and post-election case weights
anes_weights <- anes %>%
  select(V200010a, V200010b)# Post-election weights

anes_pre_post <- anes %>%
  select(c(V201001:V201658p, # Pre-election vars
           V202001:V202645n)) # Post-election vars

## Actually keep response vars in
# Move response vars to a new df
# anes_response <- anes_pre_post %>%
#   select(c(V202541a:V202547,
#            V201114:V201123,
#            V202451:V202456,
#            V201626:V201627,
#            V201639:V201640,
#            V201602:V201605x,
#            V202022:V202024))
# 
# anes_pre_post <- anes_pre_post %>%
#   select(!all_of(colnames(anes_response)))
# 
# Make list of response vars to not be removed
response_vars <- anes_pre_post %>%
  select(c(V202541a:V202547,
           V201114:V201123,
           V202451:V202456,
           V201626:V201627,
           V201639:V201640,
           V201602:V201605x,
           V202022:V202024)) %>%
  colnames()

# Remove vars with 0 variance
n_unique <- apply(anes_pre_post,
                  2,
                  function(x) length(unique(x)))

cols_to_keep <- names(n_unique)[n_unique > 1]

anes_pre_post <- anes_pre_post %>%
  select(all_of(cols_to_keep))
```


```{r}
# Get the pre-election media responses that are measured by Ad Fontes Media
adfontes <- read_csv('data/raw/adfontes.csv')
media_var_names <- adfontes$Code

anes_media <- anes %>%
  select(all_of(media_var_names))

anes_response <- anes %>%
  select(all_of(response_vars))

# Remove variables lowly correlated with media vars
media_cors <- cor(anes_pre_post %>% select(!all_of(media_var_names) &
                                           !all_of(response_vars)),
    anes_pre_post %>% select(all_of(media_var_names)),
    method = 'pearson')

media_cors_df <- media_cors %>%
  as.data.frame()

rownames(media_cors_df) <- rownames(media_cors)

max_cors <- apply(media_cors_df,
                  1,
                  max)

vars_to_remove <- names(which(max_cors < 0.1))

anes_pre_post2 <- anes_pre_post %>%
  select(!all_of(vars_to_remove))

# Print number of variables removed
length(vars_to_remove)

# Remove highly correlated variables
anes_other_vars <- anes_pre_post2 %>%
                  select(!all_of(media_var_names) &
                         !all_of(response_vars))

anes_cor <- cor(anes_other_vars,
                method = 'spearman')

anes_cor[upper.tri(anes_cor)] <- 0
diag(anes_cor) <- 0

cols_to_remove <- anes_other_vars[, apply(anes_cor,
                                         2,
                                         function(x) any(abs(x) > 0.90,
                                                         na.rm = TRUE))] %>%
  colnames()

anes_pre_post3 <- anes_pre_post2 %>%
  select(!all_of(cols_to_remove))

# Print # of columns removed
ncol(anes_pre_post2) - ncol(anes_pre_post3)

# Print final # of variables
ncol(anes_pre_post3)
```

For the media variables set any values of -9 (Refused), -5 (Interview breakoff (sufficient partial IW)), and -1 (inapplicable) to missing values (NA).

```{r set-na}
# In media vars convert -9 (Refused), -5 (Interview breakoff (sufficient partial IW)), and
# -1 (inapplicable) to NA
anes_to_impute <- anes_pre_post3 %>%
  mutate(across(everything(),
                as.numeric),
         across(all_of(media_var_names),
                ~ as.factor(as.character(ifelse(. < 0, NA, .))))) %>%
  as.data.frame()
```

```{r}
# Creat a list of NULL values the length of the number of columns for the imputation dataset
# impute_classwts <- vector(mode = 'list',
#                           length = ncol(anes_to_impute))
# 
# # Add the column names to the list for matching
# names(impute_classwts) <- colnames(anes_to_impute)
# 
# # Calculate the case weights for each factor variable
# media_classwts <- anes_to_impute %>%
#   select(where(is.factor)) %>%
#   pivot_longer(everything()) %>%
#   drop_na() %>%
#   group_by(name) %>%
#   summarize(n_samples = n(),
#             n_samples0 = sum(value == '0'),
#             n_samples1 = sum(value == '1'),
#             '0' = n_samples / (2 * n_samples0),
#             '1' = n_samples / (2 * n_samples1)) %>%
#   select(name, '0', '1')
# 
# # Add the case weights for each factor variable, leaving continuous variables as NULL
# for (ii in 1:nrow(media_classwts)) {
#   index <- which(names(impute_classwts) %in% media_classwts[ii, 'name'])
#   
#   impute_classwts[[index]] <- cbind(media_classwts[ii, '0'], 
#                                    media_classwts[ii, '1'])
# }
```

```{r}
library(mice)

n_cores <- parallel::detectCores()

anes_imputed <- futuremice(anes_to_impute,
                         m = n_cores,
                         parallelseed = 4960,
                         n.core = n_cores,
                         method = 'rf')

saveRDS(anes_imputed, "data/interim/imputed-mice-rf.RDS")
```

For each of the imputed datasets, calculate the total and average Bias and Reliability scores.

```{r}
anes_imputed %$% lm(V201630b ~ V202413) %>% pool %$% pooled

model_formula <- as.formula(paste("V202452 ~", paste(media_vars, collapse = "+")))

with(anes_imputed, expr = vglm)
```

