---
title: "Imputation with random forests"
author: "Tyler Garner"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/anes-media-bias")
```

Load in the data set and remove all of the variables with zero variance (have only a single value).  Also drop all of the variables that will be used as responses in modeling.

As 

```{r load-libs-and-data, warning = FALSE}
library(tidyverse)

anes <- read_csv("data/raw/anes_timeseries_2020_csv_20220210.csv",
                 show_col_types = FALSE)

# Keep only pre- and post-election vars and post-election case weights
anes_weights <- anes %>%
  select(V200010b)# Post-election weights

anes <- anes %>%
  select(c(V201001:V201658p, # Pre-election vars
           V202001:V202645n)) # Post-election vars

# Move response vars to a new df
anes_response <- anes %>%
  select(c(V202541a:V202547,
            V201114:V201123,
            V202451:V202456,
            V201626:V201627,
            V201639:V201640))

anes <- anes %>%
  select(!all_of(colnames(anes_response)))

# Remove vars with 0 variance
n_unique <- apply(anes,
                  2,
                  function(x) length(unique(x)))

cols_to_keep <- names(n_unique)[n_unique > 1]

anes <- anes %>%
  select(all_of(cols_to_keep))

# Get the pre-election media responses
anes_media <- anes %>%
  select(V201630a:V201630r, # TV programs 1
         V201631a:V201631r, # TV programs 2
         V201633a:V201633r, # Radio programs
         V201634a:V201634c,V201634e:V201634f, V201634h,
         V201634j:V201634q, # Websites not included in 'online newspapers'
         V201636a:V201636d) # Online newspapers
```

For the media variables set any values of -9 (Refused), -5 (Interview breakoff (sufficient partial IW)), and -1 (inapplicable) to missing values (NA).

```{r set-na}
# Get the names of all media vars of interest
anes_media_vars <- anes_media %>%
  colnames()

# In media vars convert -9 (Refused), -5 (Interview breakoff (sufficient partial IW)), and
# -1 (inapplicable) to NA
anes_to_impute <- anes %>%
  mutate(across(everything(),
                as.numeric),
         across(all_of(anes_media_vars),
                ~ as.factor(as.character(ifelse(. < 0, NA, .))))) %>%
  as.data.frame()
```

### Random forest imputation

The `missForest()` from the `missForest` library will be used to impute the missing values generated above using a random forest model. This approach has the advantage of being able to impute categorical data while handling complex interactions and nonlinear relationships. However, the algorithm has a long computation time, particularly when there are a lot of variables in the dataset.

First, the imputation method will be tested on known data to check whether the imputation on unknown data can be trusted.

```{r impute-test}
library(missForest)
library(doParallel)

# Select rows where all media variables do not contain missing values
impute_test_df <- anes_to_impute %>%
  filter(if_all(V201630a:V201636d, ~ !is.na(.)))

test_to_impute <- impute_test_df

set.seed(1)
V201630a_na <- impute_test_df %>%
  select(V201630a) %>%
  prodNA()

test_to_impute <- test_to_impute %>%
  mutate(V201630a = as.numeric(V201630a_na$V201630a) - 1,
         across(V201630b:V201630r, ~ ifelse(is.na(V201630a), NA, .) - 1),
         across(V201630a:V201630r, ~ as.factor(as.character(.))))

set.seed(2)
V201631a_na <- impute_test_df %>%
  select(V201631a) %>%
  prodNA()

test_to_impute <- test_to_impute %>%
  mutate(V201631a = as.numeric(V201631a_na$V201631a) - 1,
         across(V201631b:V201631r, ~ ifelse(is.na(V201631a), NA, .) - 1),
         across(V201631a:V201631r, ~ as.factor(as.character(.))))

set.seed(3)
V201633a_na <- impute_test_df %>%
  select(V201633a) %>%
  prodNA()

test_to_impute <- test_to_impute %>%
  mutate(V201633a = as.numeric(V201633a_na$V201633a) - 1,
         across(V201633b:V201633q, ~ ifelse(is.na(V201633a), NA, .) - 1),
         across(V201633a:V201633q, ~ as.factor(as.character(.))))

set.seed(4)
V201634a_na <- impute_test_df %>%
  select(V201634a) %>%
  prodNA()

test_to_impute <- test_to_impute %>%
  mutate(V201634a = as.numeric(V201634a_na$V201634a) - 1,
         across(c(V201634b:V201634c,V201634e:V201634f, V201634h, V201634j:V201634q),
                ~ ifelse(is.na(V201634a), NA, .) - 1),
         across(c(V201634a:V201634c,V201634e:V201634f, V201634h, V201634j:V201634q),
         ~ as.factor(as.character(.))))

set.seed(5)
V201636a_na <- impute_test_df %>%
  select(V201636a) %>%
  prodNA()

test_to_impute <- test_to_impute %>%
  mutate(V201636a = as.numeric(V201636a_na$V201636a) - 1,
         across(V201636b:V201636d, ~ ifelse(is.na(V201636a), NA, .) - 1),
         across(V201636a:V201636d, ~ as.factor(as.character(.))))

# Create parallel backend
n_cores <- parallel::detectCores()

cl <- makeForkCluster(n_cores)
registerDoParallel(cl)

set.seed(4960)
test_imputed <- test_to_impute %>%
  missForest(verbose = TRUE,
             maxiter = 10,
             # xtrue = impute_test_df, this does not work ??
             # Error in ximp[mis] - xtrue[mis] : non-numeric argument to binary operator
             parallelize = 'forests')

stopCluster(cl)

test_imputed_table <- table(impute_test_df[is.na(test_to_impute)],
      test_imputed$ximp[is.na(test_to_impute)])

test_imputed_table

sum(diag(test_imputed_table)) / sum(test_imputed_table)
```

The total accuracy predicting the missing variables was 90.5%, with 90.9% of '0's and 69.2% of '1's correctly imputed.

The function by default will run over 10 iterations and takes a sizeable amount of time with this dataset, so a parallel backend will be registered using all available cores to save computation time.

```{r impute}
library(missForest)
library(doParallel)

# Create parallel backend
n_cores <- parallel::detectCores()

cl <- makeForkCluster(n_cores)
registerDoParallel(cl)

set.seed(4960)
anes_imputed <- anes_to_impute %>%
  missForest(verbose = TRUE,
             maxiter = 10,
             parallelize = 'forests')

stopCluster(cl)

saveRDS(anes_imputed, 'data/interim/missForst-imputed.RDS')
```


```{r}
anes %>%
  filter(across(c(V201114:V201123,
                  V202451:V202455,
                  V201626:V201627,
                  V201639:V201640,
         V201630a:V201630r, # TV programs 1
         V201631a:V201631r, # TV programs 2
         V201633a:V201633r, # Radio programs
         V201634a:V201634c,V201634e:V201634f, V201634h,
         V201634j:V201634q, # Websites not included in 'online newspapers'
         V201636a:V201636d), ~ . >= 0))
```

