---
title: "Imputation with random forests"
author: "Tyler Garner"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/anes-media-bias")
```

Load in the data set and remove all of the variables with zero variance (have only a single value).  Also drop all of the variables that will be used as responses in modeling.

As 

```{r load-libs-and-data, warning = FALSE}
library(tidyverse)

anes <- read_csv("data/raw/anes_timeseries_2020_csv_20220210.csv",
                 show_col_types = FALSE)

# Keep only pre- and post-election vars and post-election case weights
anes_weights <- anes %>%
  select(V200010b)# Post-election weights

anes <- anes %>%
  select(c(V201001:V201658p, # Pre-election vars
           V202001:V202645n)) # Post-election vars

# Move response vars to a new df
anes_response <- anes %>%
  select(c(V202541a:V202547,
           V201114:V201123,
           V202451:V202456,
           V201626:V201627,
           V201639:V201640,
           V201602:V201605x,
           V202022:V202024))

anes <- anes %>%
  select(!all_of(colnames(anes_response)))

# Remove vars with 0 variance
n_unique <- apply(anes,
                  2,
                  function(x) length(unique(x)))

cols_to_keep <- names(n_unique)[n_unique > 1]

anes <- anes %>%
  select(all_of(cols_to_keep))

# Get the pre-election media responses
anes_media <- anes %>%
  select(V201630a:V201630r, # TV programs 1
         V201631a:V201631r, # TV programs 2
         V201633a:V201633r, # Radio programs
         V201634a:V201634c,V201634e:V201634f, V201634h,
         V201634j:V201634q, # Websites not included in 'online newspapers'
         V201636a:V201636d) # Online newspapers
```

For the media variables set any values of -9 (Refused), -5 (Interview breakoff (sufficient partial IW)), and -1 (inapplicable) to missing values (NA).

```{r set-na}
# Get the names of all media vars of interest
anes_media_vars <- anes_media %>%
  colnames()

# In media vars convert -9 (Refused), -5 (Interview breakoff (sufficient partial IW)), and
# -1 (inapplicable) to NA
anes_to_impute <- anes %>%
  mutate(across(everything(),
                as.numeric),
         across(all_of(anes_media_vars),
                ~ as.factor(as.character(ifelse(. < 0, NA, .))))) %>%
  as.data.frame()
```

```{r}
# Creat a list of NULL values the length of the number of columns for the imputation dataset
impute_classwts <- vector(mode = 'list', length = ncol(anes_to_impute))

# Add the column names to the list for matching
names(impute_classwts) <- colnames(anes_to_impute)

# Calculate the case weights for each factor variable
media_classwts <- anes_to_impute %>%
  select(where(is.factor)) %>%
  pivot_longer(everything()) %>%
  drop_na() %>%
  group_by(name) %>%
  summarize(n_samples = n(),
            n_samples0 = sum(value == '0'),
            n_samples1 = sum(value == '1'),
            '0' = n_samples / (2 * n_samples0),
            '1' = n_samples / (2 * n_samples1)) %>%
  select(name, '0', '1')

# Add the case weights for each factor variable, leaving continuous variables as NULL
for (ii in 1:nrow(media_classwts)) {
  index <- which(names(impute_classwts) %in% media_classwts[ii, 'name'])
  
  impute_classwts[[index]] <- cbind(media_classwts[ii, '0'], 
                                   media_classwts[ii, '1'])
}
```


### Random forest imputation

The `missForest()` from the `missForest` library will be used to impute the missing values generated above using a random forest model. This approach has the advantage of being able to impute categorical data while handling complex interactions and nonlinear relationships. However, the algorithm has a long computation time, particularly when there are a lot of variables in the dataset.

The function by default will run over 10 iterations and takes a sizeable amount of time with this dataset, so a parallel backend will be registered using all available cores to save computation time.

```{r impute eval = FALSE}
library(missForest)
library(doParallel)

# Create parallel backend
n_cores <- parallel::detectCores()

cl <- makeForkCluster(n_cores)
registerDoParallel(cl)

set.seed(4960)
anes_imputed <- anes_to_impute %>%
  missForest(verbose = TRUE,
             maxiter = 10,
             classwt = impute_classwts,
             parallelize = 'forests') # Note that parallelizing over 'variables' will cause errors

stopCluster(cl)

saveRDS(anes_imputed, 'data/interim/missForst-imputed.RDS')
```

The imputation ended at iteration 9 as the differences increased and so iteration 8 was taken as the final estimate.
The out-of-bag (OOB) proportion of falsely classified predictions is 0.08082, which can be considered equivalent to ~ 92% accuracy since the variables being imputed on are binary.


Calculate # of rows of non-missing data in responses

```{r}
anes %>%
  filter(across(c(V201114:V201123,
                  V202451:V202455,
                  V201626:V201627,
                  V201639:V201640,
         V201630a:V201630r, # TV programs 1
         V201631a:V201631r, # TV programs 2
         V201633a:V201633r, # Radio programs
         V201634a:V201634c,V201634e:V201634f, V201634h,
         V201634j:V201634q, # Websites not included in 'online newspapers'
         V201636a:V201636d), ~ . >= 0))
```



